---
title: "mode-share-comparison"
author: "jafshin"
date: "05/07/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Intro

This is code compares the mode share of trips from Simulation output to 
- Simulation input
- VISTA
- Census JTW for work trips

## Mode share comparison for All trips

### Simulation output processing

```{r reading sim outputs}

simOutputTrips <- read_delim(gzfile("data/simOutputs/10pct/CalibRun6/output_trips.csv.gz"),
                      delim=";") %>% 
  # Removing trips with zero distance
  filter(traveled_distance>0) %>%   
  filter(longest_distance_mode!="walk") %>% 
  mutate(longest_distance_mode=ifelse(longest_distance_mode=="netwalk", "walk", longest_distance_mode)) 

```

```{r}

simOutputTripsCounted <- simOutputTrips %>% 
  count(mode=longest_distance_mode) %>% 
  # Getting the pct
  mutate(pct=100*n/sum(n)) 

```

### Simulation input processing

```{r}
demandOutputDir <- "data/demand/output-Sep10-10pct/"
simInputTrips <- read_csv(paste0(demandOutputDir,"7.time/plan.csv")) %>% 
  # filtering those without mode
  filter(!is.na(ArrivingMode)) %>%
  # filtering the trips that are not a vakid trip!
  filter(!is.na(Distance)) %>% 
  # renaming bike with bicycle
  mutate(ArrivingMode=ifelse(ArrivingMode=="bike","bicycle",ArrivingMode)) 

```

```{r}
simInputTripsCounted <- simInputTrips %>% 
  # counting trips per mode
  count(mode=ArrivingMode) %>% 
  # getting the percentage 
  mutate(pct=100*n/sum(n))
```

### VISTA trips processing

```{r}

vistaTrips <- read_csv("~/ownCloud/Data/VISTA/2012-18/T_VISTA1218_V1.csv")

vistaTripsCounted <- vistaTrips %>% 
  # Filtering to valid trips
  filter(CUMDIST>0) %>% 
  # Refactoring the linkmode in VISTA
  mutate(mode=case_when(LINKMODE=="Bicycle" ~ "bicycle",
                        LINKMODE=="Vehicle Driver" ~ "car", # Check with Dhirendra/Alan what happens to car passengers?
                        LINKMODE=="Walking" ~ "walk",
                        LINKMODE%in%c("Public Bus","Train","Tram") ~ "pt",
                        TRUE ~ "other")) %>% 
  # Filtering out modes==other
  filter(mode!="other") %>% 
  # counting trips per mode
  # count(mode) %>% 
  filter(!is.na(WDTRIPWGT)) %>% 
  group_by(mode) %>% 
  summarise(n=sum(WDTRIPWGT)) %>% 
  # get the percentages
  mutate(pct=100*n/sum(n))
vistaTripsCounted
```

### Plotting them all 

```{r}

rbind(mutate(simOutputTripsCounted,source="Simulation Output"),
      mutate(simInputTripsCounted,source="Simulation "),
      mutate(vistaTripsCounted,source="VISTA Trips")
      ) %>% 
    mutate(source=factor(source,levels=c("VISTA Trips","IVN"),labels=c("SON","IVN"))) %>%   
  ggplot(aes(x=mode, y=pct, fill=source)) +
  geom_col(position="dodge") +
  geom_text(aes(label=paste0(round(pct, digits = 1),"%")), 
            position=position_dodge(width=0.9), vjust=-0.25)

```


## Mode share comparison for JTW trips

```{r Census JTW}

censusJTW <- read_csv("~/ownCloud/Data/ABS_Census/SA3 (UR) and MTW15P Method of Travel to Work (15 travel modes).csv", 
                        skip=9, ) %>%
  rename(HomeSA3=`SA3 (UR)`,
         MTW=`MTW15P Method of Travel to Work (15 travel modes)`) %>%
  dplyr::select(-Counting, -X5) %>%
  filter(!is.na(MTW)) %>% 
  filter(MTW!="Total") %>% 
  # filter(HomeSA3 %in% unique(HomeSA3)) %>% 
  mutate(mode=case_when(MTW=="Bicycle" ~ "bicycle",
                        MTW%in%c("Car, as driver") ~ "car", # Not including ,"Car, as passenger"
                        MTW%in%c("Bus","Train", "Tram") ~ "pt",
                        MTW=="Walked only" ~ "walk",
                        TRUE ~ "other")) %>% 
  filter(mode!="other") 



censusJTWCounted <- censusJTW %>% 
  group_by(mode) %>% 
  summarise(n=sum(Count)) %>% 
  mutate(pct=100*n/sum(n)) %>% 
  dplyr::select(mode, n, pct)

censusJTWCounted
```

```{r VISTA JTW}

vistaJTW <- read_csv("~/ownCloud/Data/VISTA/2012-18/JTW_VISTA1218_V1.csv") %>% 
  # Filtering to valid trips
  filter(JTWDIST>0) %>% 
  # Refactoring the linkmode in VISTA
  mutate(mode=case_when(JTWMODE=="Bicycle" ~ "bicycle",
                        JTWMODE=="Vehicle Driver" ~ "car", # Check with Dhirendra/Alan what happens to car passengers?
                        JTWMODE=="Walking" ~ "walk",
                        JTWMODE%in%c("Public Bus","Train","Tram") ~ "pt",
                        TRUE ~ "other")) %>% 
  # Filtering out modes==other
  filter(mode!="other") 

vistaJTWCounted <- vistaJTW %>% 
  # counting trips per mode
  # count(mode) %>% 
  filter(!is.na(WDJTWWGT)) %>% 
  group_by(mode) %>% 
  summarise(n=sum(WDJTWWGT)) %>% 
  # get the percentages
  mutate(pct=100*n/sum(n))

vistaJTWCounted
```


```{r Simulation input JTW}

simInputJTW <- simInputTrips %>% 
  filter(Activity=="Work") 
  
simInputJTWCounted <- simInputJTW %>% 
  # counting trips per mode
  count(mode=ArrivingMode) %>% 
  # getting the percentage 
  mutate(pct=100*n/sum(n))

simInputJTWCounted
```


```{r Simulation Output JTW}

simOutputJTW <- simOutputTrips %>% 
  filter(end_activity_type=="Work") 
  
simOutputJTWCounted <- simOutputJTW  %>% 
  # Counting number of trips with each mode
  count(mode=longest_distance_mode) %>% 
  # Getting the pct
  mutate(pct=100*n/sum(n)) 

simOutputJTWCounted
```

### Plotting JTW trips

```{r}
workTripsJoined <-rbind(mutate(simOutputJTWCounted,source="Simulation Output"),
      # mutate(simInputJTWCounted,source="Simulation Input JTW"),
      mutate(vistaJTWCounted,source="VISTA 2012-16"),
      mutate(censusJTWCounted,source="Census 2016")
      ) %>% 
  mutate(type="Mandatory trips - work")

workTripsJoined %>% 
  ggplot(aes(x=mode, y=pct, fill=source)) +
  geom_col(position="dodge") +
  geom_text(aes(label=paste0(round(pct, digits = 1),"%")), 
            position=position_dodge(width=0.9), vjust=-0.25)

```

## education trips Analysis

```{r VISTA education trips}

vistaEdu <- vistaTrips %>%
  filter(DESTPURP1%in%c("Education")) %>% 
  # Filtering to valid trips
  filter(CUMDIST>0) %>% 
  # Refactoring the linkmode in VISTA
  mutate(mode=case_when(LINKMODE=="Bicycle" ~ "bicycle",
                        LINKMODE=="Vehicle Driver" ~ "car", # Check with Dhirendra/Alan what happens to car passengers?
                        LINKMODE=="Walking" ~ "walk",
                        LINKMODE%in%c("Public Bus","Train","Tram") ~ "pt",
                        TRUE ~ "other")) %>% 
  # Filtering out modes==other
  filter(mode!="other") 

vistaEduCounted <- vistaEdu %>% 
  # counting trips per mode
  # count(mode) %>% 
  filter(!is.na(WDTRIPWGT)) %>% 
  group_by(mode) %>% 
  summarise(n=sum(WDTRIPWGT)) %>% 
  # get the percentages
  mutate(pct=100*n/sum(n))

vistaEduCounted
```

```{r Simulation input education trips}

simInputEdu <- simInputTrips %>% 
  filter(Activity=="Study") 
  
simInputEduCounted <- simInputEdu %>% 
  # counting trips per mode
  count(mode=ArrivingMode) %>% 
  # getting the percentage 
  mutate(pct=100*n/sum(n))

simInputEduCounted
```

```{r Simulation Output education trips}

simOutputEdu <- simOutputTrips %>% 
  filter(end_activity_type=="Study") 
  
simOutputEduCounted <- simOutputEdu  %>% 
  # Counting number of trips with each mode
  count(mode=longest_distance_mode) %>% 
  # Getting the pct
  mutate(pct=100*n/sum(n)) 

simOutputEduCounted
```

### Plotting education trips
```{r}

educationTripsJoined <- rbind(mutate(simOutputEduCounted,source="Simulation Output"),
      # mutate(simInputEduCounted,source="Simulation Input"),
      mutate(vistaEduCounted,source="VISTA 2012-16")
      )  %>% 
  mutate(type="Mandatory trips - education")

educationTripsJoined %>% 
  ggplot(aes(x=mode, y=pct, fill=source)) +
  geom_col(position="dodge") +
  geom_text(aes(label=paste0(round(pct, digits = 1),"%")), 
            position=position_dodge(width=0.9), vjust=-0.25)

```



## Discretionary trips Analysis

```{r VISTA Discretionary trips}

vistaDiscretionary <- vistaTrips %>%
  filter(!DESTPURP1%in%c("Education", "Work Related")) %>% 
  # Filtering to valid trips
  filter(CUMDIST>0) %>% 
  # Refactoring the linkmode in VISTA
  mutate(mode=case_when(LINKMODE=="Bicycle" ~ "bicycle",
                        LINKMODE=="Vehicle Driver" ~ "car", # Check with Dhirendra/Alan what happens to car passengers?
                        LINKMODE=="Walking" ~ "walk",
                        LINKMODE%in%c("Public Bus","Train","Tram") ~ "pt",
                        TRUE ~ "other")) %>% 
  # Filtering out modes==other
  filter(mode!="other") 

vistaDiscretionaryCounted <- vistaDiscretionary %>% 
  # counting trips per mode
  # count(mode) %>% 
  filter(!is.na(WDTRIPWGT)) %>% 
  group_by(mode) %>% 
  summarise(n=sum(WDTRIPWGT)) %>% 
  # get the percentages
  mutate(pct=100*n/sum(n))

vistaDiscretionaryCounted
```

```{r Simulation input Discretionary trips}

simInputDiscretionary <- simInputTrips %>% 
  filter(!Activity%in%c("Study","Work")) 
  
simInputDiscretionaryCounted <- simInputDiscretionary %>% 
  # counting trips per mode
  count(mode=ArrivingMode) %>% 
  # getting the percentage 
  mutate(pct=100*n/sum(n))

simInputDiscretionaryCounted
```

```{r Simulation Output Discretionary trips}

simOutputDiscretionary <- simOutputTrips %>% 
  filter(!end_activity_type%in%c("Study","Work")) 
  
simOutputDiscretionaryCounted <- simOutputDiscretionary  %>% 
  # Counting number of trips with each mode
  count(mode=longest_distance_mode) %>% 
  # Getting the pct
  mutate(pct=100*n/sum(n)) 

simOutputDiscretionaryCounted
```

### Plotting Discretionary trips
```{r}
discretionaryCountsJoined <- rbind(
  mutate(simOutputDiscretionaryCounted,source="Simulation Output"),
  # mutate(simInputDiscretionaryCounted,source="Simulation Input"),
  mutate(vistaDiscretionaryCounted,source="VISTA 2012-16")
  ) %>% 
  mutate(type="Discretionary trips")

discretionaryCountsJoined %>% 
  ggplot(aes(x=mode, y=pct, fill=source)) +
  geom_col(position="dodge") +
  geom_text(aes(label=paste0(round(pct, digits = 1),"%")), 
            position=position_dodge(width=0.9), vjust=-0.25)

```



## NonWork trips Analysis

```{r VISTA NonWork trips}

vistaNonWork <- vistaTrips %>%
  filter(DESTPURP1!="Work Related") %>% 
  # Filtering to valid trips
  filter(CUMDIST>0) %>% 
  # Refactoring the linkmode in VISTA
  mutate(mode=case_when(LINKMODE=="Bicycle" ~ "bicycle",
                        LINKMODE=="Vehicle Driver" ~ "car", # Check with Dhirendra/Alan what happens to car passengers?
                        LINKMODE=="Walking" ~ "walk",
                        LINKMODE%in%c("Public Bus","Train","Tram") ~ "pt",
                        TRUE ~ "other")) %>% 
  # Filtering out modes==other
  filter(mode!="other") 

vistaNonWorkCounted <- vistaNonWork %>% 
  # counting trips per mode
  # count(mode) %>% 
  filter(!is.na(WDTRIPWGT)) %>% 
  group_by(mode) %>% 
  summarise(n=sum(WDTRIPWGT)) %>% 
  # get the percentages
  mutate(pct=100*n/sum(n))

vistaNonWorkCounted
```

```{r Simulation input NonWork trips}

simInputNonWork <- simInputTrips %>% 
  filter(Activity!="Work") 
  
simInputNonWorkCounted <- simInputNonWork %>% 
  # counting trips per mode
  count(mode=ArrivingMode) %>% 
  # getting the percentage 
  mutate(pct=100*n/sum(n))

simInputNonWorkCounted
```

```{r Simulation Output NonWork trips}

simOutputNonWork <- simOutputTrips %>% 
  filter(end_activity_type!="Work") 
  
simOutputNonWorkCounted <- simOutputNonWork  %>% 
  # Counting number of trips with each mode
  count(mode=longest_distance_mode) %>% 
  # Getting the pct
  mutate(pct=100*n/sum(n)) 

simOutputNonWorkCounted
```

### Plotting NonWork trips
```{r}
NonWorkCountsJoined <- rbind(
  mutate(simOutputNonWorkCounted,source="Simulation Output"),
  # mutate(simInputNonWorkCounted,source="Simulation Input"),
  mutate(vistaNonWorkCounted,source="VISTA 2012-16")
  ) %>% 
  mutate(type="NonWork trips")

NonWorkCountsJoined %>% 
  ggplot(aes(x=mode, y=pct, fill=source)) +
  geom_col(position="dodge") +
  geom_text(aes(label=paste0(round(pct, digits = 1),"%")), 
            position=position_dodge(width=0.9), vjust=-0.25)

```


## Table for the paper:

```{r}
library(xtable)

allModeSharesJoined <- rbind(workTripsJoined,
educationTripsJoined,
NonWorkCountsJoined,
discretionaryCountsJoined) %>% 
  mutate(pct2=paste0(round(pct, digits = 1),"%")) %>% 
  pivot_wider(id_cols = c("type","mode"), names_from = "source", values_from="pct2") 

print(xtable(allModeSharesJoined), booktabs = TRUE)

```

