---
title: "road-traffic-comparison"
author: "jafshin"
date: "16/06/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(sf)
library(viridis)
library(ggspatial)
```


## Road traffic comparison

This document is to compare traffic volumes from real-world observations to simulation outputs.

Steps before this step:
- Running **observation-preparation.Rmd** to process real-world observations
- Running **simOutputProcessing.Rmd** to process to create traffic volume joined



Run the chunk below to process simulation outputs



## Car traffic

```{r Reading input car traffic data}

amCarObservations <- read_sf("./data/observationsJoined/carAmObsJoined2Network.sqlite") %>% 
  dplyr::select(link_id, starts_with("x")) 
pmCarObservations <- read_sf("./data/observationsJoined/carPmObsJoined2Network.sqlite") %>% 
  dplyr::select(link_id, starts_with("x"))

if (all(colnames(amCarObservations)==colnames(pmCarObservations))) {
  if (nrow(filter(amCarObservations,link_id%in%pmCarObservations$link_id))==0) {
      carObservations <- rbind(st_drop_geometry(amCarObservations), 
                               st_drop_geometry(pmCarObservations))
  }else print("There is a duplicate link_id in AM and PM data!")
}else print("Colnames in AM and PM car observations does not match!")

carSimulation <- read_sf("./data/simOutputJoined/networkLinksWithHourlyVol.sqlite", 
                         layer="car")

```


```{r restructuring observation and simulation car data}

carObservationsLong <- carObservations %>% 
  pivot_longer(cols = starts_with("x"), names_to="hour", values_to="obs_volume") %>% 
  mutate(hour=as.integer(str_remove(hour,"x"))) #%>% 
  # mutate(Source="Observation")

carSimulationFiltered <- carSimulation %>% 
  st_drop_geometry() %>% 
  dplyr::select(link_id=id, starts_with("X")) %>% 
  filter(link_id %in% carObservations$link_id)

carSimulationLong <- carSimulationFiltered %>% 
  pivot_longer(cols = starts_with("X"), names_to="hour", values_to="sim_volume") %>% 
  mutate(hour=as.double(str_remove(hour,"X"))) %>% 
  mutate(sim_volume=sim_volume*10)  %>% # Because I used 10% sample 
  mutate(hour=case_when(hour==24 ~ 0,
                        hour==25 ~ 1,
                        hour==26 ~ 2,
                        hour==27 ~ 3,
                        hour==28 ~ 4,
                        hour==29 ~ 5,
                        hour==30 ~ 6,
                        TRUE ~ hour)) %>% 
  group_by(link_id, hour) %>% 
  summarise(sim_volume=sum(sim_volume,na.rm = T))
  # mutate(Source="Simulation")

carData <- carObservationsLong %>% 
  left_join(carSimulationLong, by = c("link_id","hour"))

```

### Plotting the volumes
54686
Daily simulated car traffic for all roads

```{r}

carSimulationFiltered <- carSimulation %>% 
  filter(total_vol>0) %>% 
  mutate(volumeAdjusted=total_vol*10) %>% 
  dplyr::select(volumeAdjusted)

carBreakPoints <- pretty(carSimulationFiltered$volumeAdjusted, n = 6)

ggplot(carSimulationFiltered) +
  annotation_map_tile(type="osmgrayscale",zoom=9, alpha=0.6) +
  geom_sf(aes(size=volumeAdjusted, colour=volumeAdjusted, alpha=volumeAdjusted)) +
  # geom_sf(aes( size=n, colour=n ))  +
  scale_color_viridis(option="magma", trans="sqrt", breaks=carBreakPoints, name="Daily volume" ) +
  scale_alpha_continuous(name="Daily volume", trans="sqrt", range=c(0.05, .9), breaks=carBreakPoints) +
  scale_size_continuous(name="Daily volume", range=c(0.01, 2), breaks=carBreakPoints) +
    theme_void() + 
    guides( colour = guide_legend()) +
    theme(
      legend.position = c(0.85, 0.8),
      text = element_text(color = "#22211d"),
      plot.background = element_rect(fill = "#f5f5f2", color = NA), 
      panel.background = element_rect(fill = "#f5f5f2", color = NA), 
      legend.background = element_rect(fill = "#f5f5f2", color = NA)
    )

ggsave("~/Dropbox/Apps/Overleaf/MATSimMelbournePaper/figs/carDailyTraffic.png",
       width = 7, height = 7)
```


```{r}

carData %>% 
  # filter(link_id==11169) %>% 
  ggplot(aes(x=hour)) +
  geom_line(aes(y=obs_volume, color="observation")) + 
  geom_line(aes(y=sim_volume, color="simulation")) +
  facet_wrap("link_id")

ggsave(paste0(outputDir,"car.png"), width= 50, height = 50, units="cm")

```
Selecting two properly matched car data

```{r}
# carData %>% 
#   filter(link_id%in%c(11011, 116340)) %>% 
#   # filter(link_id==11169) %>% 
#   ggplot(aes(x=hour)) +
#   geom_line(aes(y=obs_volume, color="observation")) + 
#   geom_line(aes(y=sim_volume, color="simulation")) +
#   facet_wrap("link_id") +
#   theme(
#       legend.position = "bottom",
#       legend.title = element_blank(),
#       text = element_text(color = "#22211d"),
#       plot.background = element_rect(fill = "#f5f5f2", color = NA), 
#       panel.background = element_rect(fill = "#f5f5f2", color = NA), 
#       legend.background = element_rect(fill = "#f5f5f2", color = NA)
#     ) +
#   xlab("Hour of the day") +
#   ylab("Traffic volume")
# 
# ggsave("~/Dropbox/Apps/Overleaf/MATSimMelbournePaper/figs/carSelected.png",
#        width= 7, height = 4)
```


All car on the road traffice

```{r}

carSimHourlyAll <- carSimulationLong %>% 
  filter(!is.na(sim_volume)) %>% 
  filter(hour<24) %>% 
  group_by(hour) %>% 
  summarise(volume=sum(sim_volume)) %>%
  mutate(volume_ratio=100*volume/sum(volume)) %>% 
  mutate(source="Simulation")

carObsHourlyAll <- carObservationsLong %>% 
  group_by(hour) %>% 
  summarise(volume=sum(obs_volume)) %>% 
    mutate(volume_ratio=100*volume/sum(volume)) %>% 
  mutate(source="Observation")

rbind(carObsHourlyAll,
carSimHourlyAll) %>% 
  # filter(link_id==11169) %>% 
  ggplot(aes(x=hour,y=volume_ratio, color=source)) +
  geom_col(aes(fill=source),position = "identity" , alpha=0.2) +
  geom_line() + 
  theme(
      legend.position = "bottom",
      legend.title = element_blank(),
      text = element_text(color = "#22211d"),
      plot.background = element_rect(fill = "#f5f5f2", color = NA), 
      panel.background = element_rect(fill = "#f5f5f2", color = NA), 
      legend.background = element_rect(fill = "#f5f5f2", color = NA)
    ) +
  xlab("Hour of the day") +
  ylab("Traffic volume %")

ggsave("~/Dropbox/Apps/Overleaf/MATSimMelbournePaper/figs/carTrafficOveral.png",
       width= 7, height = 4)
```


Calculating GEH Statistic for car traffic

```{r}

carDataWithGEH <- carData %>% 
  # mutate(sim_volume=sim_volume*1.25) %>% 
  mutate(gehStat= sqrt((2*(sim_volume-obs_volume)^2)/(sim_volume+obs_volume)))

st_write(carDataWithGEH, paste0(outputDir,"calibrationData.sqlite"), 
         layer="car", delete_layer=T)
```

Plotting the GEH for car traffic

```{r}

carDataWithGEH %>% 
  # filter(link_id==11169) %>% 
  ggplot(aes(x=hour)) +
  geom_line(aes(y=gehStat)) + 
  facet_wrap("link_id")

ggsave(paste0(outputDir,"carGeh.png"), width= 50, height = 50, units="cm")

```

```{r}
carDataWithGEH %>% 
  drop_na(gehStat) %>% 
  group_by(link_id) %>% 
  summarise(avgGEH=mean(gehStat)) %>% 
  view()
```

### Using precentage error instead of GEH

```{r}

carDataWithGEH <- carData %>% 
  filter(!link_id%in%c(54473, 162636, 119121)) %>% 
  # filter(hour%in%c(6,7,8,9,10,16,17,18,19)) %>% 
  # mutate(sim_volume=sim_volume*1.25) %>% 
  mutate(gehStat= sqrt((2*(sim_volume-obs_volume)^2)/(sim_volume+obs_volume))) %>% 
  drop_na(gehStat)

carDataWithGEH %>% 
  ggplot()+
  geom_histogram(aes(x=gehStat))

{carDataWithGEH %>% filter(gehStat<=10) %>% nrow()} / nrow(carDataWithGEH)
{carDataWithGEH %>% filter(gehStat<=5) %>% nrow()} / nrow(carDataWithGEH)

carDataWithApe <- carData %>% 
  mutate(sim_volume=sim_volume*1.25) %>% 
  mutate(ape= abs(obs_volume-sim_volume)/obs_volume )

carDataWithApe %>% 
  # filter(link_id==11169) %>% 
  ggplot(aes(x=hour)) +
  geom_line(aes(y=ape)) + 
  facet_wrap("link_id")
ggsave(paste0(outputDir,"car_ape.png"), width= 50, height = 50, units="cm")

```

### Comparing hourly shares

```{r}

carDataPct.h <- carData %>% 
  mutate(obs_volume=replace_na(obs_volume,0),
         sim_volume=replace_na(sim_volume,0)) %>% 
  group_by(link_id) %>% 
  mutate(obs.pct.h=obs_volume/sum(obs_volume, na.rm = T),
         sim.pct.h=sim_volume/sum(sim_volume, na.rm = T)) 

carDataPct.h %>% 
  ggplot(aes(x=hour))+
 geom_line(aes(y=obs.pct.h, color="observation")) + 
  geom_line(aes(y=sim.pct.h, color="simulation")) +
  facet_wrap("link_id")

ggsave(paste0(outputDir,"carpct.png"), width= 50, height = 50, units="cm")

```

Hourly aggregated WAPE

```{r}
error_wape <- carDataPct.h %>% 
  filter(!link_id%in%c(54473, 162636, 119121, 112409,118506,163016,116571,163535, 152409)) %>% 
  ungroup() %>% 
  # filter(hour%in%7:18) %>% 
  group_by(hour) %>% 
  summarise(wape=100*sum(abs(obs.pct.h-sim.pct.h))/sum(obs.pct.h))

error_wape %>% 
  ggplot(aes(x=hour, y=wape)) +
  geom_line()

ggsave(paste0(outputDir,"car_wape.png"), width= 5, height = 5)

```

Aggregated

```{r}
carDataPct.h %>% 
    filter(!link_id%in%c(54473, 162636, 119121, 112409,118506,163016,116571,163535, 152409)) %>% 
  group_by(hour) %>% 
  summarise(obs.pct.h.avg=mean(obs.pct.h),
            sim.pct.h.avg=mean(sim.pct.h)) %>% 
    ggplot(aes(x=hour))+
 geom_line(aes(y=obs.pct.h.avg, color="observation")) + 
  geom_line(aes(y=sim.pct.h.avg, color="simulation"))

ggsave(paste0(outputDir,"carpctavg.png"), width= 5, height = 5)

```

## Bike traffic

```{r Reading input bike traffic data}

bikeObservations <- read_sf("./data/observationsJoined/cyclingObsJoined2Network.sqlite") %>% 
  dplyr::select(link_id, hour, obs_volume = count) %>% 
  st_drop_geometry()

bikeSimulation <- read_sf("./data/simOutputJoined/networkLinksWithHourlyVol.sqlite", 
                         layer="bike")

```


```{r restructuring observation and simulation bike data}

bikeSimulationFiltered <- bikeSimulation %>% 
  st_drop_geometry() %>% 
  dplyr::select(link_id=id, starts_with("X")) %>% 
  filter(link_id %in% bikeObservations$link_id)

bikeSimulationLong <- bikeSimulationFiltered %>% 
  pivot_longer(cols = starts_with("X"), names_to="hour", values_to="sim_volume") %>% 
  mutate(hour=as.integer(str_remove(hour,"X"))) %>% 
  mutate(sim_volume=sim_volume*10) # %>% # Because I used 10% sample
  # mutate(Source="Simulation")

bikeData <- bikeObservations %>% 
  left_join(bikeSimulationLong, by = c("link_id","hour"))

```

### Plotting the volumes

Daily simulated bike traffic for all roads

```{r}

bikeSimulationFiltered <- bikeSimulation %>% 
  filter(total_vol>0) %>% 
  mutate(volumeAdjusted=total_vol*10) %>% 
  dplyr::select(volumeAdjusted)

breakPoints <- pretty(bikeSimulationFiltered$volumeAdjusted, n = 6)

ggplot(bikeSimulationFiltered) +
  annotation_map_tile(type="osmgrayscale",zoom=9, alpha=0.6) +
  geom_sf(aes(size=volumeAdjusted, colour=volumeAdjusted, alpha=volumeAdjusted)) +
  # geom_sf(aes( size=n, colour=n ))  +
  scale_color_viridis(option="magma", trans="sqrt", breaks=breakPoints, name="Daily volume" ) +
  scale_alpha_continuous(name="Daily volume", trans="sqrt", range=c(0.1, .9), breaks=breakPoints) +
  scale_size_continuous(name="Daily volume", range=c(0.01, 2), breaks=breakPoints) +
    theme_void() + 
    guides( colour = guide_legend()) +
    theme(
      legend.position = c(0.85, 0.8),
      text = element_text(color = "#22211d"),
      plot.background = element_rect(fill = "#f5f5f2", color = NA), 
      panel.background = element_rect(fill = "#f5f5f2", color = NA), 
      legend.background = element_rect(fill = "#f5f5f2", color = NA)
    )

ggsave("~/Dropbox/Apps/Overleaf/MATSimMelbournePaper/figs/bikeDailyTraffic.png",
       width = 7, height = 7)
```


```{r}

bikeData %>% 
  # filter(link_id==11169) %>% 
  ggplot(aes(x=hour)) +
  geom_line(aes(y=obs_volume, color="observation")) + 
  geom_line(aes(y=sim_volume, color="simulation")) +
  facet_wrap("link_id")

ggsave(paste0(outputDir,"bike.png"), width= 50, height = 50, units="cm")

```

Selecting two properly matched bike data

```{r}
bikeData %>% 
  filter(link_id%in%c(191054, 446314)) %>% 
  # filter(link_id==11169) %>% 
  ggplot(aes(x=hour)) +
  geom_line(aes(y=obs_volume, color="observation")) + 
  geom_line(aes(y=sim_volume, color="simulation")) +
  facet_wrap("link_id") +
  theme(
      legend.position = "bottom",
      legend.title = element_blank(),
      text = element_text(color = "#22211d"),
      plot.background = element_rect(fill = "#f5f5f2", color = NA), 
      panel.background = element_rect(fill = "#f5f5f2", color = NA), 
      legend.background = element_rect(fill = "#f5f5f2", color = NA)
    ) +
  xlab("Hour of the day") +
  ylab("Traffic volume")

ggsave("~/Dropbox/Apps/Overleaf/MATSimMelbournePaper/figs/bikeSelected.png",
       width= 7, height = 4)
```

All bikes on the road traffice

```{r}

bikeSimHourlyAll <- bikeSimulationLong %>% 
  filter(!is.na(sim_volume)) %>% 
  filter(hour<24) %>% 
  group_by(hour) %>% 
  summarise(volume=sum(sim_volume)) %>% 
  mutate(volume_ratio =100* volume/sum(volume)) %>% 
  mutate(source="Simulation")

bikeObsHourlyAll <- bikeObservations %>% 
  group_by(hour) %>% 
  summarise(volume=sum(obs_volume)) %>% 
  mutate(volume_ratio =100* volume/sum(volume)) %>% 
  mutate(source="Observation")

rbind(bikeObsHourlyAll,
bikeSimHourlyAll) %>% 
  # filter(link_id==11169) %>% 
  ggplot(aes(x=hour,y=volume_ratio, color=source)) +
  geom_col(aes(fill=source), position = "identity" ,   alpha=0.2) +
  geom_line() + 
  theme(
      legend.position = "bottom",
      legend.title = element_blank(),
      text = element_text(color = "#22211d"),
      plot.background = element_rect(fill = "#f5f5f2", color = NA), 
      panel.background = element_rect(fill = "#f5f5f2", color = NA), 
      legend.background = element_rect(fill = "#f5f5f2", color = NA)
    ) +
  xlab("Hour of the day") +
  ylab("Traffic volume %")

ggsave("~/Dropbox/Apps/Overleaf/MATSimMelbournePaper/figs/bikeTrafficOveral.png",
       width= 7, height = 6)
```


Calculating GEH Statistic for Bike traffic

```{r}

bikeDataWithGEH <- bikeData %>% 
  mutate(gehStat= sqrt((2*(sim_volume-obs_volume)^2)/(sim_volume+obs_volume)))
st_write(bikeDataWithGEH, paste0(outputDir,"calibrationData.sqlite"), 
         layer="bike", delete_layer=T)
```

Plotting the GEH for Bike traffic

```{r}

bikeDataWithGEH %>% 
  # filter(link_id==11169) %>% 
  ggplot(aes(x=hour)) +
  geom_line(aes(y=gehStat)) + 
  facet_wrap("link_id")

ggsave(paste0(outputDir,"bikeGeh.png"), width= 50, height = 50, units="cm")

```


