---
title: "simulationOutputPorcessing"
author: "jafshin"
date: "15/06/2021"
output: html_document
params:
  parseXML: FALSE
---

## Simulation output post-processing for Calibration/validation

This code converts MATSim's event file into four different CSV files, each capturing certain types of events.
Then it processes the link enter/exit events to figure out traffic volume by hour for each link of the network.
Next, it reads MATSim's XML output network and converts it to SF objects and then adds traffic volume to each network link to be used for calibration/validation.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(ggspatial)
library(sf)
library(vroom)
library(fs)
source("./functions/convertXMLNetwork.R")
print(params$parseXML)
```

### Converting event.xml output to CSVs

```{r cleaning old files, eval=params$parseXM}
paths <- dir_ls("./data/simOutput/",glob = "*.txt" )
if (length(paths)>0) {
  walk(paths, ~file.remove(.x))
}
```


```{bash processing event XMLs, eval=params$parseXM}

#!/bin/bash
if [[ $# -eq 0 ]]; then
  events_file=./data/simOutput/output_events.xml.gz;
else
  events_file=$1
fi

# Departure and Arrival
zcat $events_file | tail -n +3 | head -n -2 |
grep "departure\|arrival" |
cut -d "\"" -f 2,4,6,8,10 |
sed -e "s/\"/,/g"|
sed -e '1i\'$'\n''time,type,person,link,legMode' | cat > ./data/simOutput/person_trip.txt

# left link and entered link
zcat $events_file | tail -n +3 | head -n -2 |
grep "left link\|entered link" |
sed -e "s/left link/left_link/g; s/entered link/entered_link/g " |
cut -d "\"" -f 2,4,6,8 |
sed -e "s/\"/,/g"|
sed -e '1i\'$'\n''time,type,link,vehicle' | cat > ./data/simOutput/trip_links.txt

# vehicle-traffic interaction
zcat $events_file | tail -n +3 | head -n -2 |
grep "vehicle leaves traffic\|vehicle enters traffic" |
sed -e "s/vehicle leaves traffic/vehicle_leaves_traffic/g;" |
sed -e "s/vehicle enters traffic/vehicle_enters_traffic/g;" |
cut -d "\"" -f 2,4,6,8,10,12 |
sed -e "s/\"/,/g"|
sed -e '1i\'$'\n''time,type,person,link,vehicle,networkMode' | cat > ./data/simOutput/vehicle_trip.txt

# Activity-startend actstart
zcat  $events_file | tail -n +3 | head -n -2 |
grep "actstart\|actend" |
cut -d "\"" -f 2,4,6,8,10 |
sed -e "s/\"/,/g"|
sed -e '1i\'$'\n''time,type,person,link,actType' | cat > ./data/simOutput/person_act.txt


```

### link volume analysis

```{r reading trip link events}

tripLinks <- vroom("./data/simOutput/trip_links.txt")
# tripLinks %>% slice_sample(prop=0.1) %>% write_delim("./data/simOutput/trip_links_small.txt")
# tripLinks <- vroom("./data/simOutput/trip_links_small.txt") 

```

Adding mode type to the trips

```{r adding mode to link events}
vehicleTrips <- read_csv("./data/simOutput/vehicle_trip.txt")

vehicleLists <- vehicleTrips %>% 
  distinct(person,vehicle,networkMode) 

tripLinksWithVehicle <- tripLinks %>% 
  left_join(vehicleLists, by = "vehicle") %>% 
  # Filtering to non-pt vehicles 
  filter(!(grepl("bus", vehicle) | grepl("train", vehicle) | grepl("tram", vehicle)))
tripLinksWithVehicle %>% count(networkMode)


```

We can either try to join to get the duration on each link, however, it will result in missing many records, instead here we only consider link exit event as our measure for count:

```{r}

exitLinksWithTime <- tripLinksWithVehicle %>% 
  filter(type=="left_link") %>% 
  # mutate(id=row_number()) %>% 
  dplyr::select(-type) %>% 
  mutate(leftHour=  floor(time / 3600)) %>% 
  mutate(leftHMS=paste(sprintf("%02d",leftHour),
                        sprintf("%02d", floor(time %% 3600 / 60)),
                        sprintf("%02d", floor(time %% 60 )),
                        sep=":"))

```

#### Aggregating exiting link event by the hour of the day

```{r plotting link exit events}

exitLinksWithTime %>% 
  group_by(leftHour) %>% 
  summarise(n=n()) %>% 
  ggplot(aes(x=leftHour, y=n)) +
  geom_col() +
  labs(title = "Aggregated exiting link event count by the hour of the day") +
  xlab("Hour of the day") +
  ylab("Link exit even count")

```


## Network level analysis

```{r processing network XML}

if (file.exists("./data/networks/networkConverted.sqlite")) {
  networkLinks <- st_read("./data/networks/networkConverted.sqlite", 
                          layer = "links", quiet=T)
}else{
  networkConverted <- convertXMLNetowrk(xmlFile = "./data/simOutput/output_network.xml",
                                        netCRS = 28355)
  networkLinks <- networkConverted[[2]]
}

```

### Road network with Daily traffic volume

Joining daily car traffic data to road network

```{r joining daily car volumes to network, eval=FALSE}

carExitLinkAggregatedDaily <- exitLinksWithTime %>% 
  filter(networkMode=="car") %>% # filter to car trips only
  mutate(id=as.character(link)) %>% 
  group_by(id) %>% 
  summarise(total_vol=n())

networkLinksWithDailyVol_car <- networkLinks %>% 
  left_join(carExitLinkAggregatedDaily, by = "id")   
st_write(networkLinksWithDailyVol_car, 
         "./data/simOutputJoined/networkLinksWithDailyVol.sqlite",
         layer="car",
         delete_layer = T)

```

plotting daily car traffic volume

```{r plotting daily car volumes on network, eval=FALSE}

# networkLinksWithDailyVol$total_vol <- st_read("networkLinksWithDailyVol.sqlite")

networkLinksWithDailyVol_car %>% 
  drop_na(total_vol) %>% 
  # filter(total_vol>10) %>% 
  ggplot() +
  # annotation_map_tile(type="osmgrayscale",zoom=9, alpha=0.6) +
  geom_sf(aes(fill=total_vol)) +
  scale_fill_viridis_c(trans = "sqrt", alpha = .8) 

```

Joining daily bike traffic data to road network

```{r joining daily bike volumes to network, eval=FALSE}

bikeExitLinkAggregatedDaily <- exitLinksWithTime %>% 
  filter(networkMode=="bicycle") %>% # filter to car trips only
  mutate(id=as.character(link)) %>% 
  group_by(id) %>% 
  summarise(total_vol=n())

networkLinksWithDailyVol_bike <- networkLinks %>% 
  left_join(bikeExitLinkAggregatedDaily, by = "id")   

st_write(networkLinksWithDailyVol_bike, 
         "./data/simOutputJoined/networkLinksWithDailyVol.sqlite",
         layer="bike",
         delete_layer = T)
```

plotting daily bike traffic volume

```{r plotting daily bike volumes on network, eval=FALSE}

# networkLinksWithDailyVol$total_vol <- st_read("networkLinksWithDailyVol.sqlite")

networkLinksWithDailyVol_bike %>% 
  drop_na(total_vol) %>% 
  # filter(total_vol>10) %>% 
  ggplot() +
  # annotation_map_tile(type="osmgrayscale",zoom=9, alpha=0.6) +
  geom_sf(aes(fill=total_vol)) +
  scale_fill_viridis_c(trans = "sqrt", alpha = .8) 

```

### Road network with Hourly traffic volume

```{r}
# tempDf <- exitLinksWithTime %>% slice_sample(n=1000)
```

Joining Hourly car traffic data to road network

```{r joining Hourly car volumes to network}

carExitLinkAggregatedHourly <- exitLinksWithTime %>%
# carExitLinkAggregatedHourly <- tempDf %>% 
  filter(networkMode=="car") %>% # filter to car trips only
  mutate(id=as.character(link)) %>% 
  group_by(id, leftHour) %>% 
  summarise(hourly_vol=n()) %>% 
  ungroup() %>% 
  pivot_wider(id_cols=id, names_from=leftHour, values_from=hourly_vol) %>% 
  mutate(total_vol = rowSums(across(where(is.numeric)),na.rm = T))
  

networkLinksWithHourlyVol_car <- networkLinks %>% 
  left_join(carExitLinkAggregatedHourly, by = "id")   
st_write(networkLinksWithHourlyVol_car, 
         "./data/simOutputJoined/networkLinksWithHourlyVol.sqlite",
         layer="car",
         delete_layer = T)

```

plotting Hourly car traffic volume

```{r plotting Hourly car volumes on network}

# networkLinksWithHourlyVol$hourly_vol <- st_read("networkLinksWithHourlyVol.sqlite")

networkLinksWithHourlyVol_car %>% 
  filter(total_vol>0) %>% 
  # filter(hourly_vol>10) %>% 
  ggplot() +
  # annotation_map_tile(type="osmgrayscale",zoom=9, alpha=0.6) +
  geom_sf(aes(fill=total_vol)) +
  scale_fill_viridis_c(trans = "sqrt", alpha = .8) 

```


Joining Hourly bike traffic data to road network

```{r joining Hourly bike volumes to network}

bikeExitLinkAggregatedHourly <- exitLinksWithTime %>%
# bikeExitLinkAggregatedHourly <- tempDf %>% 
  filter(networkMode=="bicycle") %>% # filter to car trips only
  mutate(id=as.character(link)) %>% 
  group_by(id, leftHour) %>% 
  summarise(hourly_vol=n()) %>% 
  ungroup() %>% 
  pivot_wider(id_cols=id, names_from=leftHour, values_from=hourly_vol) %>% 
  mutate(total_vol = rowSums(across(where(is.numeric)),na.rm = T))

networkLinksWithHourlyVol_bike <- networkLinks %>% 
  left_join(bikeExitLinkAggregatedHourly, by = "id")   

st_write(networkLinksWithHourlyVol_bike, 
         "./data/simOutputJoined/networkLinksWithHourlyVol.sqlite",
         layer="bike",
         delete_layer = T)
```

plotting Hourly bike traffic volume

```{r plotting Hourly bike volumes on network}

# networkLinksWithHourlyVol$hourly_vol <- st_read("networkLinksWithHourlyVol.sqlite")

networkLinksWithHourlyVol_bike %>% 
  drop_na(total_vol) %>% 
  # filter(hourly_vol>10) %>% 
  ggplot() +
  # annotation_map_tile(type="osmgrayscale",zoom=9, alpha=0.6) +
  geom_sf(aes(fill=total_vol)) +
  scale_fill_viridis_c(trans = "sqrt", alpha = .8) 

```

