---
title: "travel-distance-comparison"
author: "jafshin"
date: "05/07/2021"
output: html_document
---

## Intro

This code compares travel time and distance from Simulation output with Google Maps API
To use this code you need to have a valid Google Maps API key.
copy and paste your API key in the 'api/google.key' to be used below
Please note Google Maps API is NOT a free service - be aware of the potential costs

```{r setup, include=FALSE}
library(mapsapi)
library(sf)
library(lubridate)
library(tidyverse)
apikey <- readLines("./api/google.key")

```

```{r Reading input from simulation}


simOutputTrips <- read_delim(gzfile("data/simOutput/10pct/exp6_withMC_withSP/output_trips.csv.gz"),
                      delim=";") %>% 
  # Removing trips with zero distance
  filter(traveled_distance>0) 

# Adding time of the day and date to the trips
```

Getting the distance and time for sampled trips for Google API

```{r get dist and time}

modesMap=tibble(modes=c("pt","car","walk","bicycle"),
           gMode=c("transit","driving","walking", "bicycling"))

tripsWithDepTime <- simOutputTrips %>% 
  left_join(modesMap, by=c("longest_distance_mode"="modes")) %>% 
  mutate(depDate = paste0(Sys.Date() + 1, " ", dep_time)) %>% 
  mutate(depDate = as.POSIXct(depDate,format="%Y-%m-%d %H:%M:%OS",
                              tz = Sys.timezone())) %>% 
  filter(!is.na(depDate))
  

```

Sampling for each valid mode (adjust n)

```{r sampling the trips}

n <- 50

tripsSampled <- tripsWithDepTime %>% 
  group_by(longest_distance_mode) %>% 
  slice_sample(n = n) %>% 
  ungroup()

origins <- tripsSampled %>% 
  dplyr::select(trip_id,start_x, start_y) %>% 
  mutate(GEOMETRY=paste0("POINT(",start_x," ",start_y,")")) %>%
  st_as_sf(wkt = "GEOMETRY", crs = 28355) %>% 
  as.data.frame() %>%
  st_sf() %>% 
  st_transform(4283)

destinations <- tripsSampled %>% 
  dplyr::select(trip_id,end_x, end_y) %>% 
  mutate(GEOMETRY=paste0("POINT(",end_x," ",end_y,")")) %>%
  st_as_sf(wkt = "GEOMETRY", crs = 28355) %>% 
  as.data.frame() %>%
  st_sf() %>% 
  st_transform(4283)

```




## Assigning the Google Time and Distance to the Sample

```{r, warning=FALSE, message=FALSE}
i=1
for(i in 1:nrow(tripsSampled)){
# for(i in 1:10){
  if(i%%10 == 0) print(paste0("trip number ",i," out of ", nrow(tripsSampled), " trips"))
  route=mapsapi::mp_matrix(origins = origins[i,"GEOMETRY"],
                           destinations = destinations[i,"GEOMETRY"],
                           mode = tripsSampled$gMode[i],
                           departure_time = tripsSampled$depDate[i],
                           traffic_model="best_guess",
                           key = apikey)
  tripsSampled[i,"time_google"] <- route %>% mp_get_matrix(value = c("duration_s")) %>% as.numeric()
  tripsSampled[i,"distance_google"] <- route %>% mp_get_matrix(value = c("distance_m")) %>% as.numeric()
}
```

```{r}
tripsProcessed <- tripsSampled %>% 
  mutate(trav_time=seconds(trav_time)) %>% 
  mutate(time_google=seconds(time_google)) %>% 
  dplyr::select(person, trip_number, depDate, mode=longest_distance_mode,
                time_simulation=trav_time, distance_simulation=traveled_distance,
                time_google, distance_google) %>%
  mutate(`Travel Distance Difference`=100*(distance_simulation-distance_google)/distance_google) %>%
  mutate(`Travel Time Difference`=100*(time_simulation-time_google)/time_google) 

write_csv(tripsProcessed, "tripsProcessed_exp6_withMC_withSP_n50.csv")
tripsProcessed <- read_csv("tripsProcessed_exp6_withMC_withSP_n50.csv")
```

## Comparing the time and distance

### Time

```{r}

tripsProcessed %>% 
  filter(`Travel Time Difference`<1000) %>%
  mutate(`Travel Mode`=case_when(mode=="car" ~ "Driving",
                                 mode=="pt" ~ "PT",
                                 mode=="bicycle" ~ "Cycling",
                                 mode=="walk" ~ "Walking",
                                 TRUE ~ "NA")) %>% 
  ggplot(aes(x=`Travel Mode`, y=`Travel Time Difference`,color=`Travel Mode`, 
             fill= `Travel Mode`)) +
  geom_boxplot(alpha=0.8) +
  scale_color_brewer(palette = "Dark2") +
  scale_fill_brewer(palette = "Dark2") +
    guides( colour = guide_legend()) +
    theme(
      legend.position ="none"
    )

ggsave("~/Dropbox/Apps/Overleaf/MATSimMelbournePaper/figs/travTimeDiff.png",
      width = 5, height = 5)

```

### Distance

```{r}


tripsProcessed %>% 
  filter(`Travel Distance Difference`<500) %>%
  mutate(`Travel Mode`=case_when(mode=="car" ~ "Driving",
                                 mode=="pt" ~ "PT",
                                 mode=="bicycle" ~ "Cycling",
                                 mode=="walk" ~ "Walking",
                                 TRUE ~ "NA")) %>% 
  ggplot(aes(x=`Travel Mode`, y=`Travel Distance Difference`,color=`Travel Mode`, 
             fill= `Travel Mode`)) +
  geom_boxplot(alpha=0.8) +
  scale_color_brewer(palette = "Dark2") +
  scale_fill_brewer(palette = "Dark2") +
    guides( colour = guide_legend()) +
    theme(
      legend.position ="none"
    )

ggsave("~/Dropbox/Apps/Overleaf/MATSimMelbournePaper/figs/travDistDiff.png",
      width = 5, height = 5)


```

## Statistical test to examine the significance of difference

```{r}
#wilcox.test(formula = distance_simulation ~ distance_google, 
#                            data=tripsProcessed, paired=T, p.adjust.method="BH",conf.level=0.95)
```


